\documentclass[a4paper,english]{lipics-v2018}
\usepackage[utf8]{inputenc}
\usepackage{microtype}
\usepackage{listings}
\lstset{basicstyle=\ttfamily\small, backgroundcolor=\color{white}}
\usepackage{srcltx}
\usepackage[ttscale=.875]{libertine}
%\usepackage[colorlinks=true]{hyperref}
\usepackage{wrapfig}

\newcommand\YAMLcolonstyle{\color{red}\mdseries}
\newcommand\YAMLkeystyle{\color{black}\bfseries}
\newcommand\YAMLvaluestyle{\color{blue}\mdseries}

\hideLIPIcs

\makeatletter

\newcommand\language@yaml{yaml}

\expandafter\expandafter\expandafter\lstdefinelanguage
\expandafter{\language@yaml}
{
  keywords={true,false,null,y,n},
  keywordstyle=\color{darkgray}\bfseries,
  basicstyle=\YAMLkeystyle,                                 % assuming a key comes first
  sensitive=false,
  comment=[l]{\#},
  morecomment=[s]{/*}{*/},
  commentstyle=\color{purple}\ttfamily,
  stringstyle=\YAMLvaluestyle\ttfamily,
  moredelim=[l][\color{orange}]{\&},
  moredelim=[l][\color{magenta}]{*},
  moredelim=**[il][\YAMLcolonstyle{:}\YAMLvaluestyle]{:},   % switch to value style at :
  morestring=[b]',
  morestring=[b]",
  literate =    {---}{{\ProcessThreeDashes}}3
                {>}{{\textcolor{red}\textgreater}}1     
                {|}{{\textcolor{red}\textbar}}1 
                {\ -\ }{{\mdseries\ -\ }}3,
}

% switch to key style at EOL
\lst@AddToHook{EveryLine}{\ifx\lst@language\language@yaml\YAMLkeystyle\fi}
\makeatother

\newcommand\ProcessThreeDashes{\llap{\color{cyan}\mdseries-{-}-}}

\let\accentvec\vec
\let\spvec\vec
\let\vec\accentvec

%\addtolength{\textfloatsep}{-3mm}

\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{amsfonts}
\usepackage{verbatim}
\usepackage[textsize=tiny]{todonotes}
%\usepackage[notref,notcite]{showkeys}
%\usepackage[subtle]{savetrees}



%\sp\vspace{-1em}orem{algorithm}{Algorithm}{\bfseries}{\rmfamily}

\newcommand{\matodo}[1]{\todo[color=green!30]{#1}}
\newcommand{\alectodo}[1]{\todo[color=red!30]{#1}}
\newcommand{\eriktodo}[1]{\todo[color=blue!30]{#1}}

% Mandatory. First: Use abbreviated first/middle names. Second (only in severe cases): Use first author plus 'et. al.'
\authorrunning{M. Aumüller, E. Bernhardsson, A. Faithfull}

% Mandatory, please use full first names. LIPIcs license is "CC-BY";  http://creativecommons.org/licenses/by/3.0/
\Copyright{ Martin Aumüller, Erik Bernhardsson, Alexander Faithfull} 

% Mandatory: Please choose ACM 1998 classifications from http://www.acm.org/about/class/ccs98-html . E.g., cite as "F.1.1 Models of Computation". 
\subjclass{H.3.3 Information Search and Retrieval}

% Mandatory: Please provide 1-5 keywords
\keywords{benchmarking, nearest neighbor search, evaluation}

\usepackage{pgfplots}
\usepgfplotslibrary{groupplots,colorbrewer}
\pgfplotsset{
  legend style = {font=\ttfamily}
}
\pgfplotsset{
% initialize Set1-5:
cycle list/Set1-5,
% combine it with ’mark list*’:
cycle multiindex* list={
mark list*\nextlist
Set1-5\nextlist
},
}

\usepackage{url}

\title{ANN-Benchmarks: A Benchmarking Tool for Approximate Nearest Neighbor Algorithms%
\footnote{The research of the first and third authors has received funding from the European Research Council under the European Union's 7th Framework Programme (FP7/2007-2013) / ERC grant agreement no. 614331.
A conference version of this work was published at SISAP'17 and is available at \url{http://dx.doi.org/10.1007/978-3-319-68474-1_3}.}
}
\titlerunning{ANN-Benchmarks: A Benchmarking Tool for ANN Algorithms}

\author{Martin Aum\"{u}ller}{IT University of Copenhagen, Denmark}{maau@itu.dk}{}{}
\author{Erik Bernhardsson}{Better, Inc.}{mail@erikbern.com}{}{}
\author{Alexander Faithfull}{IT University of Copenhagen, Denmark}{alef@itu.dk}{}{}
\begin{document}
\maketitle
\begin{abstract}
    This paper describes \textsf{ANN-Benchmarks}, a tool for evaluating
    the performance of in-memory approximate nearest neighbor algorithms. It
    provides
    a standard interface for measuring the performance and quality achieved by nearest neighbor
    algorithms on different standard data sets. It
    supports several different ways of integrating $k$-NN algorithms, and
    its configuration system automatically tests a range of parameter settings
    for each algorithm.
    Algorithms are compared with respect to many different (approximate)
    quality
    measures, and adding more is easy and fast;
    the included plotting front-ends can visualise these as images, \LaTeX\
    plots, and websites with interactive plots.
    \textsf{ANN-Benchmarks} aims to provide a constantly updated overview of
    the current state of the art of $k$-NN algorithms.
    In the short term, this overview allows users to 
  	choose the correct $k$-NN algorithm and parameters for their similarity
  	search task;
	  in the longer term, algorithm designers will be able to use this overview
	  to test and refine automatic parameter tuning.
	The paper gives an overview of
    the system, evaluates the results of the benchmark,
    and points out directions for future work. Interestingly, very different approaches
	to $k$-NN search yield comparable quality-performance trade-offs. 
    The system is available at \url{http://ann-benchmarks.com}.
\end{abstract}


%\setlength{\abovecaptionskip}{0pt plus 2pt}

\section{Introduction}\label{sec:introduction}
\input{intro}
\section{Problem Definition and Quality Measures}\label{sec:problem:desc}
\input{problem}
\section{System Design}\label{sec:system:design}
\input{system}
\section{Evaluation}\label{sec:evaluation}
\input{evaluation}
\section{Conclusion \& Further Work}\label{sec:conclusion}
\input{conclusion}
\bibliographystyle{splncs03}
\bibliography{lit}


\end{document}
